---
title: "GITT in patients with aquired brain injury"
author: "Johannes Enevoldsen"
output:
  html_document:
  theme: readable

---
	
```{r global options, include=FALSE, echo=FALSE}

library(knitr)

```


```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source("main.R")

library(pander)
library(broom)

```
### Overview of patients

```{r echo=FALSE, results='asis'}
set.alignment('left', row.names = 'right')
pander(rbind(c("n", table_data_html_pt$Total),
		 c("Males", table_data_html_pt$Sex[2]),
		 c("Age - median (IQR)", table_data_html_pt$Age),
		 c("Days since injury - median (IQR)", table_data_html_pt$`Days since injury`),
		 c("FIM score - median (IQR)", table_data_html_pt$`FIM score`)
		 ))


```

### Overview of controls

```{r echo=FALSE, results='asis'}
set.alignment('left', row.names = 'right')
pander(rbind(c("n", table_data_html_cnt$Total),
		 c("Males", table_data_html_cnt$Sex[2]),
		 c("Age - median (IQR)", table_data_html_cnt$Age)
		 ))


```

#### Age ranges
Patients: `r range(df_pt$age)`

Controls: `r range(df_cnt$age)`


### Age and gender differences between patients and controls.
```{r echo=FALSE, results='markup'}
mean(df_pt$age) - mean(df_cnt$age)
var.test(age~grp, short_data)
t.test(age~grp, var.equal = TRUE, short_data)
t.test(sex=="Male"~grp, short_data)

sex_table <- table(short_data$sex, short_data$grp)
chisq.test(sex_table, correct = FALSE)
```

### FIM score
```{r}
median(df_pt$days_rtg_to_fim)
quantile(df_pt$days_rtg_to_fim)

median(df_pt$fim)
quantile(df_pt$fim)

cor.test(df_pt$fim, df_pt$age, method = "spearman")
cor.test(df_pt$fim, df_pt$gitt, method = "spearman")
qplot(age, fim, data=df_pt)
qplot(fim, gitt, data=df_pt)

```


### GITT difference between patients and controls
Welch T-test is used due to different variance in the two groups.
```{r echo=FALSE, results='markup'}
var.test(gitt~grp, data = short_data)
sd(df_cnt$gitt)
sd(df_pt$gitt)

t.test(df_cnt$gitt)
t.test(df_pt$gitt)
t.test(gitt~grp, var.equal = FALSE, short_data)
qqnorm(df_pt$gitt, main = "ABI")
qqline(df_pt$gitt)
qqnorm(df_cnt$gitt, main = "Controls")
qqline(df_cnt$gitt)




```

### Linear model of GITT by group (pt vs cnt), age and sex
```{r echo=FALSE, results='asis'}
tidy(gitt_age_sex_lm) %>% 
	cbind(confint(gitt_age_sex_lm)) %>% 
	kable()
```


### Days since injury
```{r}
plot(days_since_inj~fim, data=df_pt)
plot(days_since_inj~gitt, data=df_pt)
hist(df_pt$days_since_inj)
cor.test(df_pt$days_since_inj, df_pt$gitt, method = "spearman")
cor.test(df_pt$days_since_inj, df_pt$fim, method = "spearman")
```

### Laxative use. 

FIM is lower in laxative users

```{r}
t.test(df_pt$use_laxative == "Y", df_cnt$use_laxative == "Y")

var.test(gitt~use_laxative, data = df_pt) #Equal var

(gitt_by_lax <- t.test(gitt~use_laxative, data = df_pt, var.equal = TRUE))
-diff(gitt_by_lax$estimate) #mean Y - mean N


lapply(c("age", "days_since_inj", "fim", "fim_no_bowel"), function(x){
	formula1 <- as.formula(paste(x,"~use_laxative"))
	means <- aggregate(formula1, data = df_pt, FUN = mean)
	
	list(means,
	     diff = means[1,2]-means[2,2],
	wilcox.test(formula1, data=df_pt))
})

```

### Opioids
```{r}
t.test(df_pt$use_opioid == "Y")

var.test(gitt~use_opioid, data = df_pt) #Equal var

(gitt_by_opioid <- t.test(gitt~use_opioid, data = df_pt, var.equal = TRUE))
diff(gitt_by_opioid$estimate) 
-gitt_by_opioid$conf.int #mean Y - mean N

```

### Statins
```{r}
t.test(df_pt$use_statin)

var.test(gitt~use_statin, data = df_pt) #Equal var

(gitt_by_statin <- t.test(gitt~use_statin, data = df_pt, var.equal = TRUE))
diff(gitt_by_statin$estimate) #mean Y - mean N
-gitt_by_statin$conf.int
```

### SSRI
```{r}
t.test(df_pt$use_ssri)

var.test(gitt~use_ssri, data = df_pt) #Equal var

(gitt_by_ssri <- t.test(gitt~use_ssri, data = df_pt, var.equal = TRUE))
diff(gitt_by_ssri$estimate) #mean Y - mean N
-gitt_by_ssri$conf.int

```


### Questionnaire
```{r}
hist(as.numeric(df_pt$rtg_to_q))
range(as.numeric(df_pt$rtg_to_q), na.rm = TRUE)


df_pt$rtg_to_q %>% quantile(na.rm=TRUE)

with(df_pt, q_sum[q_sum < 10] %>% mean(na.rm = TRUE))
with(df_pt, q_sum[q_sum < 10] %>% range(na.rm = TRUE))

mean(df_pt$days_inj_to_q, na.rm=TRUE)
quantile(df_pt$days_inj_to_q, na.rm = TRUE)

```

### HRV results
```{r}
kable(hrv_table_linear) #Only for Heart rate
kable(hrv_table_log)
```

### HRV Correlation tests

Gitt does not correlate with HF power or RMSSD
```{r, echo=FALSE}
cor.test(df_pt$gitt, log(df_pt$HF.power_5)) %>% tidy() %>% kable(digits = 3, caption="Pearson: gitt, log(HF.power_5)")
cor.test(df_pt$gitt, log(df_pt$RMSSD_5)) %>% tidy() %>% kable(digits = 3, caption="Pearson: gitt, log(RMSSD_5)")
```


